Modelling low Mach number stellar hydrodynamics with MAESTROeX

### 2019-10-03 astronum_2019_MAESTROeX/astronum-maestroex.tex
minor changes

\MarginPar{I didn't think we did Zenodo yet}
New versions of the codes are released on the first of each month, and are archived on Zenodo\footnote{\url{https://zenodo.org/}}. This also provides us with DOIs (digital object identifiers), further enhancing the reproducibility of our results and ensuring the sustainability of the codes used to produce them.

@@

The massive stars that we are interested in modelling are believed to be relatively slow rotators\MarginPar{insert some citation/quantify this}, so it is likely that one of the simpler approaches outlined above should be sufficient. However, for modelling e.g.~type I X-ray bursts, it may be that a more sophisticated approach is needed. Type I X-ray bursts are produced by thermonuclear deflagrations in the liquid surface layers of fast rotating neutron stars. These stars typically rotate at frequencies of $\sim$300\textendash600 Hz, a significant fraction of the stars' breakup velocity. This produces non-negligible deformation of the star about the equator, and may cause surface flows with a significant dependence on the meriodional angle. To properly capture the effects of rotation in such a system, it may therefore be necessary to use a two-dimensional base state, with the base state pressure a function of both the radius and the meriodional angle $p_0(r)\rightarrow p_0(r, \theta)$.
\MarginPar{should meriodional be meridional (throughout text)?}

@@

\maestroex~is a new and improved version of our previous code \maestro. It is based on \amrex~(rather than \fboxlib), which allows us to exploit its powerful new solvers and make use of ongoing improvements. \maestroex~is part of the \amrexastro~suite of open source adaptive mesh refinement astrophysical hydrodynamics codes (which also includes \castro~and \nyx). New features developed in other codes in the family can therefore be easily implemented in \maestroex. As development is open, results from our simulations can be reproduced, and monthly archiving to Zenodo ensures software sustainability.\MarginPar{Zenodo?}

We're currently working on implementing rotation in \maestroex. For modelling the interiors of massive stars (which rotate relatively slowly), we may be able to get away with neglecting the centrifugal force and simply add the Coriolis force as a source term. However, this will not be sufficient for faster rotating systems (e.g.~for modelling Type I X-ray bursts on the surfaces of millisecond pulsars), so we are also exploring other methods including modelling the centrifugal force as an effective pressure term, rewriting the base state as a function of the effective gravitational potential and using a two-dimensional base state which is also a function of the meriodional angle.

>>>>

% \MarginPar{I didn't think we did Zenodo yet}
New versions of the codes are released on the first of each month. In the case of \castro, these versions are then archived on Zenodo\footnote{\url{https://zenodo.org/}}. This also provides us with DOIs (digital object identifiers), further enhancing the reproducibility of our results and ensuring the sustainability of the codes used to produce them. We intend to extend this archiving procedure to the other \amrexastro~codes in the near future.

@@

The massive stars that we are interested in modelling are believed to be relatively slow rotators\MarginPar{insert some citation/quantify this}, so it is likely that one of the simpler approaches outlined above should be sufficient. However, for modelling e.g.~type I X-ray bursts, it may be that a more sophisticated approach is needed. Type I X-ray bursts are produced by thermonuclear deflagrations in the liquid surface layers of fast rotating neutron stars. These stars typically rotate at frequencies of $\sim$300\textendash600 Hz, a significant fraction of the stars' breakup velocity. This produces non-negligible deformation of the star about the equator, and may cause surface flows with a significant dependence on the meridional angle. To properly capture the effects of rotation in such a system, it may therefore be necessary to use a two-dimensional base state, with the base state pressure a function of both the radius and the meridional angle $p_0(r)\rightarrow p_0(r, \theta)$.

@@

\maestroex~is a new and improved version of our previous code \maestro. It is based on \amrex~(rather than \fboxlib), which allows us to exploit its powerful new solvers and make use of ongoing improvements. \maestroex~is part of the \amrexastro~suite of open source adaptive mesh refinement astrophysical hydrodynamics codes (which also includes \castro~and \nyx). New features developed in other codes in the family can therefore be easily implemented in \maestroex. As development is open, results from our simulations can be reproduced, and monthly versioning ensures software sustainability.

We're currently working on implementing rotation in \maestroex. For modelling the interiors of massive stars (which rotate relatively slowly), we may be able to get away with neglecting the centrifugal force and simply add the Coriolis force as a source term. However, this will not be sufficient for faster rotating systems (e.g.~for modelling Type I X-ray bursts on the surfaces of millisecond pulsars), so we are also exploring other methods including modelling the centrifugal force as an effective pressure term, rewriting the base state as a function of the effective gravitational potential and using a two-dimensional base state which is also a function of the meridional angle.
